---
title: "&nbsp;"
output: github_document
---

[![logo](logo.svg)](https://uclacovidbehindbars.org/)

# behindbarstools

[![Travis build status](https://travis-ci.com/uclalawcovid19behindbars/behindbarstools.svg?branch=master)](https://travis-ci.com/uclalawcovid19behindbars/behindbarstools)

## Overview
`behindbarstools` is an R package with the set of data tools used by the [UCLA Law COVID-19 Behind Bars Project](https://uclacovidbehindbars.org/) – a data project that collects and reports facility-level data on COVID-19 in jails, prisons, and other carceral facilities. `behindbarstools` includes a variety of functions to help pull, clean, wrangle, and visualize our data. 

**Warning:** This package is actively under development. 

## Installation 

```{r, message=FALSE, warning=FALSE} 
# Install directly from GitHub 
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
```

## Usage Examples 

### Reading Data 
The `read_scrape_data()` function can be used to load our data. 

`behindbarstools` also includes functions to more easily load related data from other organizations including [the Vera Institute's Jail Population Data](https://github.com/vera-institute/jail-population-data) through `read_vera_pop()` and [the Department of Homeland Security's Homeland Infrastructure Foundation-Level Data](https://hifld-geoplatform.opendata.arcgis.com/datasets/prison-boundaries/data) through `read_hifld_data()`.  

```{r read, message=FALSE, warning=FALSE}
library(behindbarstools)

# Pull latest data
latest_scraped <- read_scrape_data()

# Pull historical scraped data for California 
scraped_CA <- read_scrape_data(all_dates = TRUE, 
                               state = "California")
```

### Processing Data 

The majority of the functions in `behindbarstools` help standardize our ETL and data cleaning process. This includes functions to help with the following: 

* Cleaning facility names, e.g.`clean_fac_col_txt()`, `clean_facility_name()`
* Coalescing data from various sources, e.g. `coalesce_with_warnings()`, `group_by_coalesce()`
* Enforcing data validation, e.g. `is_valid_state()`, `is_federal()`) 
* Standardizing our data scraping infrastructure, e.g. `ExtractTable()`, `wayback_archives()`, `get_src_by_attr()` 

See [our package documentation](https://github.com/uclalawcovid19behindbars/behindbarstools/tree/master/man) for more information and examples for each function.  

### Visualizing Data 

`behindbarstools` also includes functions to create data visualizations. This includes a custom `ggplot2` theme called `theme_behindbars()` that incorporates our team's style guide. All plotting functions return [`ggplot` objects](https://ggplot2.tidyverse.org/reference/ggplot.html), making it easy to customize and add additional layers. 

```{r plot-fac, message=FALSE, warning=FALSE, fig.height=4, fig.width=9, dpi=500}
# Plot cumulative COVID-19 cases in the Los Angeles Jails over the past 30 days  
plot_fac_trend(fac_name = "Los Angeles Jails", 
               state = "California", 
               metric = "Residents.Confirmed", 
               plot_days = 30, 
               auto_label = TRUE) + 
    theme_behindbars(base_size = 14) + 
    ggplot2::ylim(3500, 4000) + 
    ggplot2::theme(legend.position = "none")
```

```{r plot-spikes, message=FALSE, warning=FALSE, fig.height=5, fig.width=9, dpi=500}
# Plot the 3 facilities with the largest recent spikes in active COVID-19 cases  
plot_recent_fac_increases(metric = "Residents.Active",
                          plot_days = 60, 
                          num_fac = 3, 
                          auto_label = TRUE) + 
    theme_behindbars(base_size = 14) + 
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1), 
                   plot.tag.position = c(0.80, 0.05))
```
